---
title: "Take 2"
author: "Kyoko"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**Here I have all of the TCGA methylation data downloaded already**
**Need to download the Clinical data, clean it up, merge it (if necessary)**


```{r}
library(TCGAbiolinks)
```


# Try pulling down TCGA clinical data

```{r}
query <- GDCquery(project = "TCGA-HNSC", 
                  data.category = "Clinical",
                  data.type = "Clinical Supplement", 
                  data.format = "BCR Biotab")
GDCdownload(query)
clinical.BCRtab.all <- GDCprepare(query)
names(clinical.BCRtab.all)
```


```{r}
View(clinical.BCRtab.all$clinical_nte_hnsc)
View(clinical.BCRtab.all$clinical_patient_hnsc)
View(clinical.BCRtab.all$clinical_follow_up_v1.0_hnsc)
View(clinical.BCRtab.all$clinical_drug_hnsc)
View(clinical.BCRtab.all$clinical_radiation_hnsc)
View(clinical.BCRtab.all$clinical_omf_v4.0_hnsc)
```

Appears to be 461 samples
^Apparently the methylation dataset is n=580 so that needs to be subset already to 461.

Maybe combine the 4:
patient_hnsc
follow-up4.8 (n=461); unique rows is 346; if multiple follow ups, take the last follow up date?
drug(n=355)
radiation(n=469)

pharmaceutical_therapy_drug_name is kinda tricky to parse because the same drug can be coded differently
but "chemotherapy" column has less number of levels so may be easier to start with




Drop the first two rows of all the tables
```{r}

functdropfirsttwo<-function(x){
  y<-x[-c(1:2),]
  return (y)
}

clinical.BCRtab.all2<-lapply(clinical.BCRtab.all, functdropfirsttwo)

```

Change all of the stuff in brackets [Not Available] etc to NA
```{r}
# Not available, not applicable, unknown, not evaluated are all stored within [] and turned into NA
turntoNA <- function(x) {
  a <- grep("\\[", x)
  vec <- rep(TRUE, length(x))
  vec[a] <- NA
  x <- x[vec]
  return(x)
}
```

```{r}
clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc<-apply(clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc, 2,
                                                         function(x) turntoNA(x))  %>% as.data.frame()
clinical.BCRtab.all2$clinical_radiation_hnsc<-apply(clinical.BCRtab.all2$clinical_radiation_hnsc, 2, 
                                                    function(x) turntoNA(x))  %>% as.data.frame()
clinical.BCRtab.all2$clinical_patient_hnsc<-apply(clinical.BCRtab.all2$clinical_patient_hnsc, 2, 
                                                  function(x) turntoNA(x))  %>% as.data.frame()
clinical.BCRtab.all2$clinical_drug_hnsc<-apply(clinical.BCRtab.all2$clinical_drug_hnsc, 2, 
                                               function(x) turntoNA(x))  %>% as.data.frame()
```

Add prefix to beginning of names to make the source identifiable:
colnames(df) <-paste("2FactorData",colnames(df),sep="-")

```{r}
colnames(clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc) <-
  paste("fu48",colnames(clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc),sep="-")
colnames(clinical.BCRtab.all2$clinical_radiation_hnsc) <-
  paste("rad",colnames(clinical.BCRtab.all2$clinical_radiation_hnsc),sep="-")
colnames(clinical.BCRtab.all2$clinical_patient_hnsc) <-
  paste("patient",colnames(clinical.BCRtab.all2$clinical_patient_hnsc),sep="-")
colnames(clinical.BCRtab.all2$clinical_drug_hnsc) <-
  paste("drug",colnames(clinical.BCRtab.all2$clinical_drug_hnsc),sep="-")
```


Trying to make each one unique to the "bcr_patient_barcode"
```{r}
patient<-clinical.BCRtab.all2$clinical_patient_hnsc
rad<-clinical.BCRtab.all2$clinical_radiation_hnsc
drug<-clinical.BCRtab.all2$clinical_drug_hnsc
fu48<-clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc
```


```{r}
length(unique(patient$`patient-bcr_patient_uuid`))
length(unique(rad$`rad-bcr_patient_uuid`))
length(unique(drug$`drug-bcr_patient_uuid`))
length(unique(fu48$`fu48-bcr_patient_uuid`))
```

rad (314), drug(186), and fu48(344) need to have unique IDs per row

Radiation for primary tumor field prioritized over other regions:
```{r}
# Extract rad therapy best response (relevant column)

rad2<-rad[,c("rad-bcr_patient_barcode","rad-treatment_best_response","rad-radiation_therapy_site")]

rad3<-distinct(rad2)

rad3$`rad-radiation_therapy_site` %>% table()

rad4<-rad3[which(rad3$`rad-radiation_therapy_site`=="Primary Tumor Field"),]

unique(rad4$`rad-bcr_patient_barcode`) %>% length()


radfinal<-rad4
```

```{r}
rad[rad$`rad-bcr_patient_barcode`=='TCGA-CN-A498',]
```

drug reduce to 186 or fewer
```{r}
drug2<-drug[,c("drug-bcr_patient_barcode","drug-pharmaceutical_therapy_type")] %>%distinct() 

drug2$chemoyes<-ifelse(drug2$`drug-pharmaceutical_therapy_type` =="Chemotherapy", "Y", NA)

drug3<-drug2[,c(1,3)] %>% drop_na() %>% distinct()

drug3$`drug-bcr_patient_barcode` %>% length()

drugfinal<-drug3
```

fu48
reduce to 344

prioritize columns 

fu48-bcr_patient_barcode
"fu48-form_completion_date"
fu48-vital_status
fu48-death_days_to
fu48-tumor_status
fu48-treatment_outcome_first_course
fu48-tobacco_smokeless_use_at_dx

```{r}
which(is.na(fu48$`fu48-treatment_outcome_first_course`))

test<-fu48[which(is.na(fu48$`fu48-treatment_outcome_first_course`)),c("fu48-bcr_patient_barcode","fu48-form_completion_date","fu48-treatment_outcome_first_course")]

interestedincols<-c("fu48-bcr_patient_barcode",
"fu48-form_completion_date",
"fu48-vital_status",
"fu48-death_days_to",
"fu48-tumor_status",
"fu48-treatment_outcome_first_course",
"fu48-tobacco_smokeless_use_at_dx")

fu48_2<-fu48[,interestedincols] %>% distinct()

```


We could split this table into "prioritized last followup date" using date on "fu48-form_completion_date" and "prioritized first followup date" to get unique patient IDs:

# try left join instead of merge
x<-pt %>% left_join (fu, by="bcr_patient_barcode") %>% left_join(drug3, by="bcr_patient_barcode")
x$birthday<-as.Date(x$form_completion_date)+as.numeric(x$birth_days_to)
x$form_completion_date<-as.Date(x$form_completion_date)
x$form_completion_date_fu<-as.Date(x$form_completion_date_fu)
x$age=as.period(interval(start = x$birthday, end = x$form_completion_date))@year
x$age_fu=as.period(interval(start = x$birthday, end = x$form_completion_date_fu))@year

elapsed.time<-x$form_completion_date %--% x$form_completion_date_fu
x$years_til_fu<-as.duration(elapsed.time)/dyears(1)
x$months_til_fu<-as.duration(elapsed.time)/dmonths(1)
x$days_til_fu<-as.duration(elapsed.time)/ddays(1)

# drop values where the form_completion_date and form_completion_date_fu are the same
# or form_completion_date_fu is earlier than form_completion_date
x2<-x%>%filter(years_til_fu>0)

# only keep the follow up that was furthest out
x3<-x2 %>% group_by(bcr_patient_barcode) %>% filter(years_til_fu == max(years_til_fu))

```{r}
# prioritize the last followup date:
fu48_2$combined<-paste(fu48_2$`fu48-bcr_patient_barcode`,fu48_2$`fu48-form_completion_date`,sep="_") 

fu48_2$combined %>% unique() %>% length()

fu48_2[order(fu48_2$combined, decreasing = FALSE),]

```

Set some sort of reference date that is much earlier than any of the form completion dates, count the number of days elapsed since that date, and take the min() and max() out of any duplicates.

```{r}
# prioritize the first followup date:

min(fu48_2$`fu48-form_completion_date`)
```

Set 2012 Jan 01 as the arbitrary reference date:

# only keep the follow up that was furthest out
x3<-x2 %>% group_by(bcr_patient_barcode) %>% filter(years_til_fu == max(years_til_fu))

```{r}
# days elapsed since reference date

fu48_2$`fu48-form_completion_date` - as.Date("2012-01-01")

fu48_2$daysinceref<-as.numeric(as.Date(fu48_2$`fu48-form_completion_date`)- as.Date("2012-01-01"), units="days")

fu48_max<-fu48_2 %>% group_by(`fu48-bcr_patient_barcode`) %>% dplyr::filter(daysinceref == max(daysinceref))

fu48_min<-fu48_2 %>% group_by(`fu48-bcr_patient_barcode`) %>% dplyr::filter(daysinceref == min(daysinceref))

```


All left join to clinical.BCRtab.all2$clinical_patient_hnsc dataset

```{r}
leftjoin1<-patient %>% left_join(fu48_max, by=c("patient-bcr_patient_barcode"="fu48-bcr_patient_barcode"))

leftjoin2<-leftjoin1 %>% left_join(radfinal, by=c("patient-bcr_patient_barcode"="rad-bcr_patient_barcode"))

leftjoin3<-leftjoin2 %>% left_join(drugfinal, by=c("patient-bcr_patient_barcode"="drug-bcr_patient_barcode"))
```


```{r}
write.csv(leftjoin3, file="clinicaltablesmall.csv")
```



