---
title: "EDA take 2"
author: "Kyoko Yamaguchi"
date: "2/28/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Loading in libraries, data.
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(Rtsne)
library(umap)
library(ggplot2)
```

```{r}
# Loading in partially cleaned tables
clincsv<-read.csv(file="clinicaltablesmall.csv", row.names = 1)
cpgcsv<-read.csv(file="hnsc_cpg1000.csv", row.names = 1)
```

```{r, include=FALSE}
rownames(clincsv)<-clincsv$patient.bcr_patient_barcode #change the rownames for clinical table

ptnames<-colnames(cpgcsv)
splitted<-strsplit(ptnames, split="\\.")

joinednames<-rep(NA, length(splitted))
for (i in 1: length(splitted)){
  joinednames[i]<-paste(splitted[[i]][1], splitted[[i]][2], splitted[[i]][3], sep="-")
}

# We see that the joinednames are NOT all unique
length(unique(joinednames))

# Exploring what the fourth element of the list stores
v<-rep(NA, length(splitted))

for (i in 1:length(splitted)) {
 v[i]<- splitted[[i]][4]
}
table(v)


# We are keeping only the patient samples marked "01A" or "01B" since these are the primary tumor samples
v2<-rep(NA, length(splitted))

for (i in 1:length(splitted)){
  if(splitted[[i]][4]=="01A"|splitted[[i]][4]=="01B"){
    v2[i]<-"include"
  }
  else{
    v2[i]<-"exclude"
  }
}


splittedsubset<-splitted[which(v2=="include")]

newjoinednames<-rep(NA, length(splittedsubset))
for (i in 1: length(splittedsubset)){
  newjoinednames[i]<-paste(splittedsubset[[i]][1], splittedsubset[[i]][2], splittedsubset[[i]][3], sep="-")
}

intersect(newjoinednames, rownames(clincsv)) %>% length() #sanity check


# the indexes for "include" are the ones we should keep
newcpgcsv<-cpgcsv[,which(v2=="include")]
newcpgcsvT<-t(newcpgcsv)
newcpgcsvT<-as.data.frame(newcpgcsvT)
rownames(newcpgcsvT)<-newjoinednames

# order the rownames based on alphabetical order of one of them to make sure they're the same order
o1<-order(rownames(clincsv))
sortedclin<-clincsv[o1,]
o2<-order(rownames(newcpgcsvT))
sortedcpg<-newcpgcsvT[o2,]
match(rownames(sortedclin), rownames(sortedcpg)) # sanity check
```

After finishing up the cleaning, the clinical table that we originally read in, "clinicaltablesmall.csv", looks like this: 

```{r}
sortedclin[1:5,5:10]
```

And the CPG table, "hnsc_cpg1000.csv", looks like this:

```{r}
sortedcpg[1:5,1:5]
```

```{r, include=FALSE, eval=FALSE}
# exporting out the cleaned tables
saveRDS(sortedclin, file="hnsc_clin_sorted.RDS")
saveRDS(sortedcpg, file="hnsc_cpg1000_sorted.RDS")
```


# Missing data

The current matrix of patients vs. CPGs does not contain any missing values. There was missing data in the original matrix of patients vs. CPGs, but the missing data was 1) kicked out if the CPG had greater than 30% missing values, and 2) the missing values for any remaining CPGs were imputed via mean imputation. The code for the mean imputation and filtering is here for reference. The Rmarkdown could not be knitted from the original table because the file size of the original table of patients vs unfiltered 400,000+ CPGs was > 1GB and not possible to share over cloud to the rest of the group. This matrix was further filtered to include only the top 1000 most variably expressed CPGs (using standard deviation). 

```{r, eval=FALSE}
# Drop CPGs that have greater than X% NAs in the row (as in, X% or more of patients have "NA" values for these CGs)
mycutoff<-0.3
cutoffrowna<-mycutoff*ncol(RAW)
numberofna<-which(rowSums(is.na(RAW))/cutoffrowna > mycutoff)
RAW_cutoffna<-RAW[-numberofna,]
dim(RAW_cutoffna)

# Impute the rows that have X% or less missing
meanimputation <- function(x) {x[is.na(x)] <- mean(x, na.rm=TRUE); return(x)}
RAW_cutoffna_impute  <- as.data.frame(apply(RAW_cutoffna, 1, meanimputation ))
RAW_cutoffna_impute<-t(RAW_cutoffna_impute)

# Keep top 1000 most variably expressed CPGs
cutoff_top1000<-min(head(sort(rowSds(RAW_cutoffna_impute), decreasing=TRUE),1000))
keepers_top1000<-which(rowSds(RAW_cutoffna_impute) >= cutoff_top1000)
RAW_cutoffna_impute_top1000<-RAW_cutoffna_impute[keepers_top1000,]
dim(RAW_cutoffna_impute_top1000)
write.csv(RAW_cutoffna_impute_top1000, "hnsc_cpg1000.csv")
```

Additionally, I realized this matrix had 52 duplicate patient IDs, which have been removed because it did not make sense for these to be in the dataset, since they correspond to code "06" and "11", which stand for "metastatic" and "solid tissue normal", respectively. We are only keeping the samples labeled "01" which stands for "primary solid tumor". (Reference: https://gdc.cancer.gov/resources-tcga-users/tcga-code-tables/sample-type-codes)

```{r}
table(v)
```

There are still missing data in the matrix of patients vs. clinical information, which we can use as a table of response variables. For your reference, the code below shows how certain interesting values from the original clinical tables from TCGA were hand-selected and wrangled to generate the table "clinicaltablesmall.csv". Since some tables contained multiple entries for the same patient (ex. if they came back for more than one follow up appointment), I prioritized keeping one account over another to make the final clinical table have only one entry per "bcr_patient_barcode". For example, for the follow up table, in the case there were two or more follow up entries, I have prioritized the FINAL follow up date.

```{r, eval=FALSE}
patient<-clinical.BCRtab.all2$clinical_patient_hnsc
rad<-clinical.BCRtab.all2$clinical_radiation_hnsc
drug<-clinical.BCRtab.all2$clinical_drug_hnsc
fu48<-clinical.BCRtab.all2$clinical_follow_up_v4.8_hnsc

# Wrangling the radiation therapy table
rad2<-rad[,c("rad-bcr_patient_barcode","rad-treatment_best_response","rad-radiation_therapy_site")]
rad3<-distinct(rad2)
rad3$`rad-radiation_therapy_site` %>% table()
rad4<-rad3[which(rad3$`rad-radiation_therapy_site`=="Primary Tumor Field"),]
unique(rad4$`rad-bcr_patient_barcode`) %>% length()
radfinal<-rad4

# Wrangling of drug therapy table
drug2<-drug[,c("drug-bcr_patient_barcode","drug-pharmaceutical_therapy_type")] %>%distinct() 
drug2$chemoyes<-ifelse(drug2$`drug-pharmaceutical_therapy_type` =="Chemotherapy", "Y", NA)
drug3<-drug2[,c(1,3)] %>% drop_na() %>% distinct()
drug3$`drug-bcr_patient_barcode` %>% length()
drugfinal<-drug3

# Wrangling the follow up table
interestedincols<-c("fu48-bcr_patient_barcode",
"fu48-form_completion_date",
"fu48-vital_status",
"fu48-death_days_to",
"fu48-tumor_status",
"fu48-treatment_outcome_first_course",
"fu48-tobacco_smokeless_use_at_dx")
fu48_2<-fu48[,interestedincols] %>% distinct()

# Calculate days elapsed since reference date of January 1, 2012, which is a date that is earlier than any of the dates that the followup form was completed, yielding a positive number
fu48_2$`fu48-form_completion_date` - as.Date("2012-01-01")
fu48_2$daysinceref<-as.numeric(as.Date(fu48_2$`fu48-form_completion_date`)- as.Date("2012-01-01"), units="days")
fu48_max<-fu48_2 %>% group_by(`fu48-bcr_patient_barcode`) %>% dplyr::filter(daysinceref == max(daysinceref))

# left join all tables to to clinical.BCRtab.all2$clinical_patient_hnsc dataset
leftjoin1<-patient %>% left_join(fu48_max, by=c("patient-bcr_patient_barcode"="fu48-bcr_patient_barcode"))
leftjoin2<-leftjoin1 %>% left_join(radfinal, by=c("patient-bcr_patient_barcode"="rad-bcr_patient_barcode"))
leftjoin3<-leftjoin2 %>% left_join(drugfinal, by=c("patient-bcr_patient_barcode"="drug-bcr_patient_barcode"))

write.csv(leftjoin3, file="clinicaltablesmall.csv")
```


This is a visualization of the "missingness" of each column in the clinical table:

```{r, fig.height=6, fig.width=20}
isna<-apply(sortedclin, 2, function(x) sum(is.na(x)))
par(mar=c(20, 4, 1, 1))
barplot(isna, las=2)
abline(h=528)

```

Some variables are entirely missing, and will not be used in the analysis. Since race and gender are variables commonly associated with disease severity and could be prognostic, it will be interesting to look at these and see if they coincide with any of the clusters found when we perform unsupervised clustering on the CPG table.


# Sampling units of variables and distributions

The sampling units for the CPG table are methylation beta values, which range from 0 to 1.

```{r}
set.seed(123)
randompick<-sample(ncol(newcpgcsvT),4)

par(mfrow=c(2,2))
for (i in 1:length(randompick)){
  hist(newcpgcsvT[,i], main=colnames(newcpgcsvT)[i])
}
```

The distribution of the beta values are quite skewed, and although log transformation is generally recommended for skewed data, it does not seem to make much of a difference, and in fact, may make thinks look even more skewed:

```{r}
logtransf<-log(newcpgcsvT)

par(mfrow=c(2,2))
for (i in 1:length(randompick)){
  hist(logtransf[,i], main=colnames(logtransf)[i])
}
```

Certain clustering algorithms do perform well even with non-Gaussian data, so we could consider many clustering algorithms s.t. we are considering groups of patients with the outcome instead of individuals.

# Key figures for later use in analysis

Key figures would be the plot of dimensions reduced by methods such as principal component analysis (PCA), T-distributed Stochastic Neighbor Embedding (TSNE), and Uniform Manifold APproximation (UMAP). Each dot represents a patient, and the dimensions are TSNE dimension 1 and 2.
Note that traditional PCA using princomp() was attempted, but failed to do spectral decomposition on our matrix (n>p). 

```{r}
# TSNE
set.seed(123)
cpgtsne<-Rtsne(sortedcpg)
plot(cpgtsne$Y, xlab="TSNE dimension 1", ylab="TSNE dimension 2")
```

```{r}
# UMAP
set.seed(123)
cpgumap<-umap(sortedcpg)
plot(cpgumap[["layout"]], xlab="UMAP dimension 1", ylab="UMAP dimension 2")
```

Although TSNE and UMAP are not clustering algorithm, we can visually see that some patients do cluster together. I have seen that common clustering algorithms like K-means and partioning around medioids (PAM) tends to give clusters that do not agree with what the human brain sees from plots of reduced dimension 1 vs reduced dimension 2.
