---
title: "Untitled"
author: "Kyoko"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**I'm trying to download files from TCGA using TCGA biolinks and other stuff**

This section describes the analysis of DNA methylation using the Bioconductor package TCGAbiolinks (Colaprico et al. 2016). 

Apparently, CpG probes used for sample identification/tracking, and those on the X and Y chromosomes were removed. Any probe with at least 1 NA were removed.

https://www.bioconductor.org/packages/devel/workflows/vignettes/TCGAWorkflow/inst/doc/TCGAWorkflow.html#Epigenetic_analysis

# Try pulling down TCGA epigenetics data

```{r, eval=FALSE}
BiocManager::install("TCGAWorkflow")
BiocManager::install("TCGAWorkflowData")
```

```{r, warning=FALSE, message=FALSE}
library(TCGAWorkflowData)
library(TCGAbiolinks)
library(SummarizedExperiment)
```

Is this a more recent version of the vignette?:
https://www.bioconductor.org/packages/release/workflows/vignettes/TCGAWorkflow/inst/doc/TCGAWorkflow.html

**I will try this for TCGA-DLBC which has only 48 samples?**

Message "Downloading data for project TCGA-DLBC
GDCdownload will download 48 files. A total of 1.02109315 GB"

Error "Error in GDCdownload.aux(server, manifest.aux, name.aux, path) : 
  There was an error in the download process (we might had a connection problem with GDC server).
Please run this function it again.
Try using method = `client` or setting files.per.chunk to a small number."

When I set n=47 it seems to work fine?

```{r}
# Samples
samples_dlbcl <- matchedMetExp("TCGA-DLBC", n = 47) #if null return all


samples <- c(samples_dlbcl)

myquery1 <- GDCquery(
  project = c("TCGA-DLBC"),
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450",
  legacy = TRUE, 
  barcode = samples
)
GDCdownload(myquery1, directory = "C:/Users/yama16/Desktop/20210927/downloads")
met_dlbcl <- GDCprepare(myquery1, save = FALSE)
```

The below coded gives an error in matchedMetExp "Sorry! There is no result for your query. Please check in GDC the data available or if there is no error in your query."
```{r}
# Samples
samples_aml <- matchedMetExp("TCGA-LAML") #if null return all

samples <- c(samples_aml)

myquery2 <- GDCquery(
  project = c("TCGA-LAML"),
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450",
  legacy = TRUE, 
  barcode = samples
)
GDCdownload(myquery2, directory = "C:/Users/yama16/Desktop/20210927/downloads")
met_aml <- GDCprepare(myquery2, save = FALSE)
```

**Try to download all the GBM (glioblastoma) methylation data by just calling the experiment name "TCGA-GBM"**

```{r}
# Samples
samples_gbm <- matchedMetExp("TCGA-GBM", n = NULL) #if null return all

length(samples_gbm  )
```
^samples_gbm contains only 62 IDs; shouldn't it be more than that?

https://portal.gdc.cancer.gov/projects/TCGA-GBM
^n=423

```{r}
samples <- c(samples_gbm)

myquery3 <- GDCquery(
  project = c("TCGA-GBM"),
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450",
  legacy = TRUE, 
  barcode = samples
)
GDCdownload(myquery3, directory = "C:/Users/yama16/Desktop/20210927/downloads")
met_gbm <- GDCprepare(myquery3, save = FALSE)
```

**Maybe I can get all 423 samples if I use the explicit barcode method?**

I can manually go to portal.gdc.cancer.gov and download the .TSVs with barcodes corresponding to the GBC study:
```{r}
barcodeslist<-c("TCGA-74-6573-01A-12D-1844-05",
"TCGA-76-6661-01B-11D-1844-05",
"TCGA-76-6280-01A-21D-1844-05",
"TCGA-28-5204-01A-01D-1481-05",
"TCGA-26-5134-01A-01D-1481-05",
"TCGA-76-6283-01A-11D-1844-05",
"TCGA-28-5220-01A-01D-1481-05",
"TCGA-06-0210-01A-01D-A45W-05",
"TCGA-06-6390-01A-11D-1697-05",
"TCGA-06-6698-01A-11D-1844-05",
"TCGA-76-6664-01A-11D-1844-05",
"TCGA-06-5416-01A-01D-1481-05",
"TCGA-06-5408-01A-01D-1697-05",
"TCGA-76-6285-01A-11D-1697-05",
"TCGA-74-6575-01A-11D-1844-05",
"TCGA-19-A6J4-01A-11D-A33U-05",
"TCGA-06-5418-01A-01D-1481-05",
"TCGA-06-5859-01A-01D-1697-05",
"TCGA-06-6693-01A-11D-1844-05",
"TCGA-19-0957-01C-01D-A45W-05",
"TCGA-14-0781-01B-01D-1697-05",
"TCGA-74-6584-01A-11D-1844-05",
"TCGA-15-1444-01A-02D-1697-05",
"TCGA-06-6697-01A-11D-1844-05",
"TCGA-06-6391-01A-11D-1697-05",
"TCGA-76-6286-01A-11D-1844-05",
"TCGA-76-4931-01A-01D-1481-05",
"TCGA-28-5209-01A-01D-1481-05",
"TCGA-06-0221-01A-01D-A45W-05",
"TCGA-19-5952-01A-11D-1697-05",
"TCGA-32-1979-01A-01D-1697-05",
"TCGA-28-5214-01A-01D-1481-05",
"TCGA-76-4927-01A-01D-1481-05",
"TCGA-06-5414-01A-01D-1481-05",
"TCGA-06-0152-01A-02D-A45W-05",
"TCGA-28-6450-01A-11D-1697-05",
"TCGA-06-AABW-11A-31D-A368-05",
"TCGA-76-4935-01A-01D-1481-05",
"TCGA-14-1402-01A-01D-A45W-05",
"TCGA-19-5960-01A-11D-1697-05",
"TCGA-06-6389-01A-11D-1697-05",
"TCGA-06-5415-01A-01D-1481-05",
"TCGA-14-0736-02A-01D-2004-05",
"TCGA-76-6282-01A-11D-1697-05",
"TCGA-28-5219-01A-01D-1481-05",
"TCGA-19-4065-02A-11D-2004-05",
"TCGA-19-5956-01A-11D-1697-05",
"TCGA-19-5959-01A-11D-1697-05",
"TCGA-4W-AA9R-01A-11D-A392-05",
"TCGA-26-6174-01A-21D-1844-05",
"TCGA-19-5947-01A-11D-1697-05",
"TCGA-74-6577-01A-11D-1844-05",
"TCGA-12-5299-01A-02D-1481-05",
"TCGA-26-5136-01B-01D-1481-05",
"TCGA-81-5910-01A-11D-1697-05",
"TCGA-76-4934-01A-01D-1481-05",
"TCGA-06-0152-02A-01D-2004-05",
"TCGA-OX-A56R-01A-11D-A33U-05",
"TCGA-14-0736-01A-01D-A45W-05",
"TCGA-28-5207-01A-01D-1481-05",
"TCGA-19-A60I-01A-12D-A33U-05",
"TCGA-06-0211-02A-02D-2004-05",
"TCGA-RR-A6KB-01A-12D-A33U-05",
"TCGA-19-5951-01A-11D-1697-05",
"TCGA-74-6581-01A-11D-1844-05",
"TCGA-06-5411-01A-01D-1697-05",
"TCGA-28-2501-01A-01D-1697-05",
"TCGA-26-5132-01A-01D-1481-05",
"TCGA-26-6173-01A-11D-1844-05",
"TCGA-06-0650-01A-02D-1697-05",
"TCGA-28-5208-01A-01D-1481-05",
"TCGA-06-0171-01A-02D-A45W-05",
"TCGA-76-6657-01A-11D-1844-05",
"TCGA-32-5222-01A-01D-1481-05",
"TCGA-19-A6J5-01A-21D-A33U-05",
"TCGA-RR-A6KC-01A-31D-A33U-05",
"TCGA-14-0862-01B-01D-1844-05",
"TCGA-19-4065-01A-01D-2004-05",
"TCGA-19-1389-02A-21D-2004-05",
"TCGA-28-5216-01A-01D-1481-05",
"TCGA-06-5412-01A-01D-1697-05",
"TCGA-87-5896-01A-01D-1697-05",
"TCGA-32-1980-01A-01D-1697-05",
"TCGA-14-1402-02A-01D-2004-05",
"TCGA-06-0190-02A-01D-2004-05",
"TCGA-19-5955-01A-11D-1697-05",
"TCGA-76-6193-01A-11D-1697-05",
"TCGA-06-0211-01A-01D-A45W-05",
"TCGA-76-4929-01A-01D-1481-05",
"TCGA-76-4932-01A-01D-1481-05",
"TCGA-76-6663-01A-11D-1844-05",
"TCGA-14-0740-01B-01D-1844-05",
"TCGA-28-2510-01A-01D-1697-05",
"TCGA-81-5911-01A-12D-1844-05",
"TCGA-14-1034-02B-01D-2004-05",
"TCGA-06-A5U1-01A-11D-A33U-05",
"TCGA-19-5954-01A-11D-1697-05",
"TCGA-74-6578-01A-11D-1844-05",
"TCGA-28-5211-01C-11D-1844-05",
"TCGA-26-5135-01A-01D-1481-05",
"TCGA-19-5958-01A-11D-1697-05",
"TCGA-76-4925-01A-01D-1481-05",
"TCGA-12-5295-01A-01D-1481-05",
"TCGA-06-6699-01A-11D-1844-05",
"TCGA-06-A7TK-01A-21D-A392-05",
"TCGA-06-A6S1-01A-11D-A33U-05",
"TCGA-RR-A6KA-01A-21D-A33U-05",
"TCGA-76-6192-01A-11D-1697-05",
"TCGA-06-6388-01A-12D-1844-05",
"TCGA-06-6700-01A-12D-1844-05",
"TCGA-06-A6S0-01A-11D-A33U-05",
"TCGA-06-1804-01A-01D-1697-05",
"TCGA-26-5133-01A-01D-1481-05",
"TCGA-14-1395-01B-11D-1844-05",
"TCGA-06-6701-01A-11D-1844-05",
"TCGA-06-1806-01A-02D-1844-05",
"TCGA-06-6695-01A-11D-1844-05",
"TCGA-14-1450-01B-01D-1844-05",
"TCGA-06-0190-01A-01D-A45W-05",
"TCGA-76-6191-01A-12D-1697-05",
"TCGA-26-A7UX-01B-11D-A392-05",
"TCGA-28-5215-01A-01D-1481-05",
"TCGA-06-5858-01A-01D-1697-05",
"TCGA-41-5651-01A-01D-1697-05",
"TCGA-12-5301-01A-01D-1481-05",
"TCGA-28-5213-01A-01D-1481-05",
"TCGA-06-0125-01A-01D-A45W-05",
"TCGA-19-5950-01A-11D-1697-05",
"TCGA-06-5410-01A-01D-1697-05",
"TCGA-06-5856-01A-01D-1697-05",
"TCGA-06-6694-01A-12D-1844-05",
"TCGA-26-5139-01A-01D-1481-05",
"TCGA-14-1043-01B-11D-1844-05",
"TCGA-06-0125-02A-11D-2004-05",
"TCGA-28-5218-01A-01D-1481-05",
"TCGA-19-1389-01A-01D-A45W-05",
"TCGA-76-4928-01B-01D-1481-05",
"TCGA-06-A5U0-01A-11D-A33U-05",
"TCGA-06-A7TL-01A-11D-A392-05",
"TCGA-74-6573-11A-02D-1844-05",
"TCGA-06-5413-01A-01D-1697-05",
"TCGA-76-6656-01A-11D-1844-05",
"TCGA-19-5953-01B-12D-1844-05",
"TCGA-06-0221-02A-11D-2004-05",
"TCGA-76-4926-01B-01D-1481-05",
"TCGA-26-1442-01A-01D-1697-05",
"TCGA-4W-AA9T-01A-11D-A392-05",
"TCGA-19-0957-02A-11D-2004-05",
"TCGA-06-5417-01A-01D-1481-05",
"TCGA-41-6646-01A-11D-1844-05",
"TCGA-4W-AA9S-01A-11D-A392-05",
"TCGA-76-6662-01A-11D-1844-05",
"TCGA-06-0171-02A-11D-2004-05",
"TCGA-06-0210-02A-01D-2004-05",
"TCGA-76-6660-01A-11D-1844-05"
)

```


Downloading files from 155 samples and putting them into 1 matrix: gbmDNAmet450k.rda

```{r}
query.met.gbm <- GDCquery(
  project = "TCGA-GBM", 
  legacy = TRUE,
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450", 
  barcode = barcodeslist
)
GDCdownload(query.met.gbm)

met.gbm.450 <- GDCprepare(
  query = query.met.gbm,
  save = TRUE, 
  save.filename = "gbmDNAmet450k.rda",
  summarizedExperiment = TRUE
)
```

```{r}
df_gbm<-met.gbm.450@assays@data@listData[[1]]
```





```{r}

```






**Apparently head and neck cancer has the second most methylation data**


```{r}
samples_hnsc <- matchedMetExp("TCGA-HNSC", n = NULL) #if null return all

length(samples_hnsc)
```
^There is apparently 546 samples?

```{r}
query.met.hnsc <- GDCquery(
  project = "TCGA-HNSC", 
  legacy = TRUE,
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450", 
  barcode = barcodeslist
)
GDCdownload(query.met.hnsc )

met.hnsc.450 <- GDCprepare(
  query = query.met.hnsc,
  save = TRUE, 
  save.filename = "hnscDNAmet450k.rda",
  summarizedExperiment = TRUE
)
```

TCGA-P3-A6SX-01A-11D-A34K-05
^This worked!
```{r}
barcodeslist<-c("TCGA-P3-A6SX-01A-11D-A34K-05")

query.met.test <- GDCquery(
  project = "TCGA-HNSC", 
  legacy = TRUE,
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450", 
  barcode = barcodeslist
)
GDCdownload(query.met.test)
```

**Manually add the IDs from the error message into the query**
This is for head and neck cancer

```{r}
barcodeslist<-c(
  "TCGA-P3-A6SX-01A-11D-A34K-05",
"TCGA-DQ-5629-01A-01D-1871-05",
"TCGA-CR-5248-01A-01D-2014-05",
"TCGA-DQ-5630-01A-01D-1871-05",
"TCGA-CV-A45T-01A-11D-A24I-05",
"TCGA-DQ-7591-01A-11D-2079-05",
"TCGA-CV-5432-01A-02D-1684-05",
"TCGA-CQ-A4CG-01A-11D-A265-05",
"TCGA-CV-A45R-01A-11D-A24I-05",
"TCGA-BA-5152-01A-02D-1871-05",
"TCGA-CN-5374-01A-01D-1433-05",
"TCGA-BA-5555-01A-01D-1511-05",
"TCGA-CN-6018-01A-11D-1684-05",
"TCGA-CV-6938-11A-01D-1913-05",
"TCGA-UF-A719-01A-12D-A34K-05",
"TCGA-BA-A6DF-01A-11D-A30F-05",
"TCGA-CV-7568-01A-11D-2230-05",
"TCGA-CR-7373-01A-11D-2014-05",
"TCGA-CN-A6V6-01A-12D-A34K-05",
"TCGA-BB-7866-01A-11D-2230-05",
"TCGA-CV-5443-01A-01D-1511-05",
"TCGA-T3-A92M-01A-31D-A392-05",
"TCGA-CQ-7067-01A-11D-2230-05",
"TCGA-QK-A64Z-01A-11D-A30F-05",
"TCGA-CV-A45P-01A-11D-A24I-05",
"TCGA-CV-7411-01A-11D-2079-05",
"TCGA-CV-A6JU-01A-11D-A31M-05",
"TCGA-CV-7406-01A-11D-2079-05",
"TCGA-IQ-A61K-01A-11D-A30F-05",
"TCGA-CV-A45V-01A-21D-A254-05",
"TCGA-CQ-6223-01A-11D-1913-05",
"TCGA-HD-8314-01A-11D-2398-05",
"TCGA-CR-6482-01A-11D-1871-05",
"TCGA-CV-7432-01A-11D-2130-05",
"TCGA-CV-5436-01A-01D-1511-05",
"TCGA-MZ-A7D7-01A-21D-A34K-05",
"TCGA-IQ-A61L-01A-11D-A30F-05",
"TCGA-CQ-5330-01A-01D-1684-05",
"TCGA-CV-5966-01A-11D-1684-05",
"TCGA-CR-7380-01A-11D-2014-05",
"TCGA-CV-7095-01A-21D-2014-05",
"TCGA-CV-7438-01A-21D-2130-05",
"TCGA-CR-7391-01A-11D-2014-05",
"TCGA-CV-7437-01A-21D-2130-05",
"TCGA-CQ-6225-01A-11D-1913-05",
"TCGA-CV-5435-01A-01D-1684-05",
"TCGA-CN-5356-01A-01D-1433-05",
"TCGA-CN-A6V1-01A-12D-A34K-05",
"TCGA-CQ-6219-01A-11D-1913-05",
"TCGA-BA-A6DE-01A-22D-A31M-05",
"TCGA-CV-6939-11A-01D-1913-05",
"TCGA-CV-7263-01A-11D-2014-05",
"TCGA-F7-A623-01A-11D-A28S-05",
"TCGA-D6-A4ZB-01A-11D-A254-05",
"TCGA-CV-5431-11A-01D-1511-05",
"TCGA-CQ-A4C6-01A-11D-A254-05",
"TCGA-UF-A7JS-01A-11D-A34K-05",
"TCGA-BA-5557-01A-01D-1511-05",
"TCGA-CQ-5323-01A-01D-1684-05",
"TCGA-CQ-5326-01A-01D-1871-05",
"TCGA-DQ-5624-01A-01D-1871-05",
"TCGA-D6-A6EN-01A-11D-A31M-05",
"TCGA-P3-A6T4-01A-11D-A34K-05",
"TCGA-BA-6871-01A-11D-1871-05",
"TCGA-CV-A461-01A-41D-A265-05",
"TCGA-IQ-7632-01A-11D-2079-05",
"TCGA-CQ-A4CI-01A-11D-A265-05",
"TCGA-F7-A620-01A-11D-A28S-05",
"TCGA-BA-A6DD-01A-12D-A31M-05",
"TCGA-CV-6962-01A-11D-1913-05",
"TCGA-CV-7248-01A-11D-2014-05",
"TCGA-BA-6872-01A-11D-1871-05",
"TCGA-BA-A6DG-01A-21D-A30F-05",
"TCGA-F7-8298-01A-11D-2398-05",
"TCGA-CR-6477-01A-11D-1871-05",
"TCGA-CQ-5325-01A-01D-1684-05",
"TCGA-CR-6480-01A-11D-1871-05",
"TCGA-CN-4728-01A-01D-1433-05",
"TCGA-DQ-7596-01A-11D-2230-05",
"TCGA-CV-5434-01A-01D-1684-05",
"TCGA-CN-6013-01A-11D-1684-05",
"TCGA-BA-A6DB-01A-11D-A30F-05",
"TCGA-CR-7377-01A-11D-2014-05",
"TCGA-CR-6487-01A-11D-1871-05",
"TCGA-CR-5247-01A-01D-2014-05",
"TCGA-CR-7390-01A-11D-2014-05",
"TCGA-HD-A6I0-01A-11D-A31M-05",
"TCGA-CV-7414-01A-11D-2079-05",
"TCGA-HD-8634-01A-11D-2398-05",
"TCGA-BA-5558-01A-01D-1511-05",
"TCGA-CV-7415-01A-11D-2079-05",
"TCGA-CR-7371-01A-11D-2014-05",
"TCGA-D6-8568-01A-11D-2398-05",
"TCGA-CN-A49A-01A-11D-A24I-05",
"TCGA-DQ-7592-01A-11D-2079-05",
"TCGA-CV-5441-01A-01D-1511-05",
"TCGA-HD-A634-01A-11D-A28S-05",
"TCGA-CR-6491-01A-11D-1871-05",
"TCGA-DQ-7594-01A-11D-2230-05",
"TCGA-P3-A6T0-01A-12D-A34K-05",
"TCGA-CV-6959-11A-01D-1913-05",
"TCGA-CV-5966-11A-01D-1684-05",
"TCGA-T3-A92N-01A-11D-A392-05",
"TCGA-CV-5976-11A-01D-1684-05",
"TCGA-CV-6441-01A-11D-1684-05",
"TCGA-T2-A6X0-01A-11D-A34K-05",
"TCGA-UF-A71E-01A-31D-A34K-05",
"TCGA-F7-A50J-01A-21D-A28S-05",
"TCGA-CV-6943-01A-11D-1913-05",
"TCGA-CN-4729-01A-01D-1433-05",
"TCGA-MT-A7BN-01A-12D-A34K-05",
"TCGA-CX-7086-01A-11D-2079-05",
"TCGA-CV-6433-11A-01D-1684-05",
"TCGA-CV-6936-11A-01D-1913-05",
"TCGA-CV-6933-01A-11D-1913-05",
"TCGA-BA-6873-01A-11D-1871-05",
"TCGA-CN-6022-01A-21D-1684-05",
"TCGA-QK-A6VC-01A-23D-A34K-05",
"TCGA-CV-7245-01A-11D-2014-05",
"TCGA-CV-6953-11A-01D-1913-05",
"TCGA-CV-5443-11A-01D-1511-05",
"TCGA-T2-A6X2-01A-12D-A34K-05",
"TCGA-CQ-5327-01A-01D-1684-05",
"TCGA-CV-7253-01A-11D-2014-05",
"TCGA-CN-5361-01A-01D-1433-05",
"TCGA-CV-5979-01A-11D-1684-05",
"TCGA-CV-6960-01A-41D-2014-05",
"TCGA-4P-AA8J-01A-11D-A392-05",
"TCGA-CN-4725-01A-01D-1433-05",
"TCGA-CV-7433-01A-11D-2130-05",
"TCGA-D6-6827-01A-11D-1913-05",
"TCGA-CR-5243-01A-01D-1511-05",
"TCGA-CN-6989-01A-11D-1913-05",
"TCGA-UF-A7JH-01A-21D-A34K-05",
"TCGA-CN-6024-01A-11D-1684-05",
"TCGA-CN-A641-01A-11D-A30F-05",
"TCGA-CV-7090-01A-11D-2014-05",
"TCGA-CV-5977-11A-01D-1684-05",
"TCGA-HD-7229-01A-11D-2014-05",
"TCGA-D6-6826-01A-11D-1913-05",
"TCGA-CV-7261-01A-11D-2014-05",
"TCGA-IQ-A61O-01A-11D-A30F-05",
"TCGA-BB-7862-01A-21D-2230-05",
"TCGA-QK-A652-01A-11D-A30F-05",
"TCGA-CV-6953-01A-11D-1913-05",
"TCGA-CN-A6V3-01A-12D-A34K-05",
"TCGA-IQ-A6SH-01A-12D-A34K-05",
"TCGA-CN-6019-01A-11D-1684-05",
"TCGA-CQ-5332-01A-01D-1684-05",
"TCGA-BA-4075-01A-01D-1433-05",
"TCGA-CR-7365-01A-11D-2014-05",
"TCGA-CR-6481-01A-11D-1871-05",
"TCGA-WA-A7GZ-01A-11D-A34K-05",
"TCGA-CV-A463-01A-11D-A265-05",
"TCGA-CV-7425-01A-11D-2079-05",
"TCGA-BA-A6DA-01A-31D-A31M-05",
"TCGA-CQ-A4CH-01A-11D-A265-05",
"TCGA-CV-7409-01A-31D-2230-05",
"TCGA-CV-5978-01A-11D-1684-05",
"TCGA-CQ-6218-01A-11D-1913-05",
"TCGA-CN-A63U-01A-11D-A30F-05",
"TCGA-CV-7254-01A-11D-2014-05",
"TCGA-IQ-A61J-01A-11D-A30F-05",
"TCGA-CR-7394-01A-11D-2014-05",
"TCGA-CV-7103-11A-01D-2014-05",
"TCGA-CR-6492-01A-12D-2079-05",
"TCGA-CV-7099-01A-41D-2014-05",
"TCGA-QK-A8ZA-01A-11D-A392-05",
"TCGA-UF-A7JJ-01A-11D-A34K-05",
"TCGA-CN-4723-01A-01D-1433-05",
"TCGA-P3-A5Q6-01A-11D-A28S-05",
"TCGA-MT-A51W-01A-21D-A265-05",
"TCGA-QK-A8Z9-01B-11D-A392-05",
"TCGA-P3-A5Q5-01A-11D-A28S-05",
"TCGA-BA-7269-01A-11D-2014-05",
"TCGA-BB-8596-01A-11D-2398-05",
"TCGA-CV-6954-11A-01D-1913-05",
"TCGA-CR-7376-01A-11D-2130-05",
"TCGA-CV-A460-01A-21D-A254-05",
"TCGA-CV-5444-01A-02D-1511-05",
"TCGA-CQ-A4CA-01A-11D-A254-05",
"TCGA-KU-A66T-01A-11D-A30F-05",
"TCGA-D6-A74Q-01A-11D-A34K-05",
"TCGA-BB-4223-01A-01D-1433-05",
"TCGA-CV-7235-11A-01D-2014-05",
"TCGA-BA-4078-01A-01D-1433-05",
"TCGA-UF-A7JF-01A-11D-A34K-05",
"TCGA-DQ-7588-01A-11D-2079-05",
"TCGA-CN-5365-01A-01D-1433-05",
"TCGA-TN-A7HJ-01A-12D-A34K-05",
"TCGA-CR-6493-01A-11D-1871-05",
"TCGA-IQ-A61E-01A-22D-A30F-05",
"TCGA-BA-A4IF-01A-11D-A265-05",
"TCGA-CV-6934-11A-01D-1913-05",
"TCGA-BA-A8YP-01A-11D-A392-05",
"TCGA-CV-6937-01A-11D-2014-05",
"TCGA-CR-6473-01A-11D-1871-05",
"TCGA-P3-A5QF-01A-11D-A28S-05",
"TCGA-CQ-5329-01A-01D-1684-05",
"TCGA-CN-5373-01A-01D-1433-05",
"TCGA-BB-4224-01A-01D-1433-05",
"TCGA-CV-6436-01A-11D-1684-05",
"TCGA-CN-4734-01A-01D-1433-05",
"TCGA-CQ-7071-01A-12D-A30F-05",
"TCGA-CV-A465-01A-11D-A265-05",
"TCGA-CN-4737-01A-01D-1433-05",
"TCGA-MT-A67D-01A-31D-A30F-05",
"TCGA-CR-7364-01A-11D-2014-05",
"TCGA-IQ-7630-01A-11D-2079-05",
"TCGA-CV-6942-01A-21D-2014-05",
"TCGA-MT-A51X-01A-11D-A265-05",
"TCGA-CV-7238-11A-01D-2014-05",
"TCGA-CN-5370-01A-01D-2014-05",
"TCGA-CN-4741-01A-01D-1433-05",
"TCGA-CR-7399-01A-11D-2014-05",
"TCGA-BA-5153-01A-01D-1433-05",
"TCGA-CR-7368-01A-11D-2130-05",
"TCGA-CQ-5331-01A-02D-1871-05",
"TCGA-BA-6870-01A-11D-1871-05",
"TCGA-CV-5444-11A-01D-1511-05",
"TCGA-QK-A6IH-01A-11D-A31M-05",
"TCGA-CQ-6227-01A-11D-1913-05",
"TCGA-CN-6997-01A-11D-2014-05",
"TCGA-CR-7397-01A-11D-2014-05",
"TCGA-MT-A67F-01A-11D-A30F-05",
"TCGA-CV-6945-01A-11D-1913-05",
"TCGA-CN-5369-01A-01D-1433-05",
"TCGA-RS-A6TO-01A-32D-A34K-05",
"TCGA-CV-6956-01A-21D-2064-05",
"TCGA-CR-7369-01A-11D-2130-05",
"TCGA-CV-5971-01A-11D-1684-05",
"TCGA-CV-6950-01A-11D-1913-05",
"TCGA-CV-5430-11B-01D-1684-05",
"TCGA-C9-A480-01A-12D-A24I-05",
"TCGA-F7-7848-01A-11D-2130-05",
"TCGA-CN-4740-01A-01D-1433-05",
"TCGA-CV-5431-01A-01D-1511-05",
"TCGA-CV-7250-11A-01D-2014-05",
"TCGA-CV-A6K1-01A-11D-A31M-05",
"TCGA-BA-A4IG-01A-11D-A265-05",
"TCGA-H7-8501-01A-11D-2398-05",
"TCGA-CV-A45O-01A-21D-A24I-05",
"TCGA-CV-7178-01A-21D-2014-05",
"TCGA-CN-A640-01A-21D-A30F-05",
"TCGA-BB-A5HZ-01A-21D-A28S-05",
"TCGA-CV-7100-01A-11D-2014-05",
"TCGA-IQ-A61H-01A-11D-A30F-05",
"TCGA-CQ-7063-01A-11D-2398-05",
"TCGA-CV-6952-11A-01D-1913-05",
"TCGA-CV-A6K2-01A-11D-A31M-05",
"TCGA-CV-5441-11A-01D-1511-05",
"TCGA-CV-7178-11A-01D-2014-05",
"TCGA-BA-A4II-01A-11D-A265-05",
"TCGA-H7-A6C5-01A-11D-A30F-05",
"TCGA-CV-7263-11A-01D-2014-05",
"TCGA-CV-A6JD-01A-11D-A31M-05",
"TCGA-CN-6988-01A-11D-1913-05",
"TCGA-CQ-7072-01A-21D-A30F-05",
"TCGA-CV-7103-01A-21D-2014-05",
"TCGA-CR-7393-01A-11D-2014-05",
"TCGA-DQ-7589-01A-11D-2230-05",
"TCGA-CV-6436-11A-01D-1684-05",
"TCGA-CN-4738-01A-02D-1511-05",
"TCGA-HD-7753-01A-11D-2079-05",
"TCGA-RS-A6TP-01A-12D-A34K-05",
"TCGA-CV-5442-11A-01D-1511-05",
"TCGA-CQ-5334-01A-01D-1684-05",
"TCGA-CV-5977-01A-11D-1684-05",
"TCGA-CV-6003-11A-01D-1684-05",
"TCGA-QK-A8Z8-01A-11D-A392-05",
"TCGA-D6-6517-01A-11D-1871-05",
"TCGA-QK-A6VB-01A-12D-A34K-05",
"TCGA-CR-7401-01A-11D-2014-05",
"TCGA-CV-A45Y-01A-11D-A254-05",
"TCGA-CN-4730-01A-01D-1433-05",
"TCGA-BB-A6UM-01A-12D-A34K-05",
"TCGA-CN-6020-01A-11D-1684-05",
"TCGA-CR-7367-01A-11D-2014-05",
"TCGA-IQ-7631-01A-11D-2079-05",
"TCGA-CQ-6229-01A-11D-1913-05",
"TCGA-CV-7255-01A-11D-2014-05",
"TCGA-CV-7238-01A-11D-2014-05",
"TCGA-P3-A6T2-01A-11D-A34K-05",
"TCGA-CV-A464-01A-11D-A265-05",
"TCGA-CR-7374-01A-11D-2014-05",
"TCGA-CN-4742-01A-02D-1511-05",
"TCGA-F7-A61S-01A-11D-A28S-05",
"TCGA-CV-7413-01A-11D-2079-05",
"TCGA-CV-7440-01A-11D-2130-05",
"TCGA-CN-4739-01A-02D-1511-05",
"TCGA-CR-6467-01A-11D-1871-05",
"TCGA-CQ-6224-01A-11D-1913-05",
"TCGA-BB-A6UO-01A-12D-A34K-05",
"TCGA-CN-A498-01A-11D-A24I-05",
"TCGA-UF-A7JD-01A-11D-A34K-05",
"TCGA-MT-A67A-01A-11D-A30F-05",
"TCGA-CV-7102-01A-11D-2014-05",
"TCGA-QK-AA3J-01A-11D-A392-05",
"TCGA-CV-7243-01A-11D-2014-05",
"TCGA-CV-7097-01A-11D-2014-05",
"TCGA-CV-7091-01A-11D-2014-05",
"TCGA-DQ-7595-01A-11D-2230-05",
"TCGA-CV-7089-01A-11D-2014-05",
"TCGA-CV-A6JE-01A-11D-A31M-05",
"TCGA-CV-7429-01A-11D-2130-05",
"TCGA-CR-5250-01A-01D-1511-05",
"TCGA-CR-7398-01A-11D-2014-05",
"TCGA-CV-6943-11A-01D-1913-05",
"TCGA-CN-A497-01A-11D-A24I-05",
"TCGA-CV-5978-11A-01D-1684-05",
"TCGA-F7-A61V-01A-11D-A28S-05",
"TCGA-CV-5973-11A-01D-1684-05",
"TCGA-CV-5970-01A-11D-1684-05",
"TCGA-BA-A4IH-01A-11D-A265-05",
"TCGA-BB-A5HU-01A-11D-A28S-05",
"TCGA-CN-6023-01A-11D-1684-05",
"TCGA-CX-A4AQ-01A-11D-A254-05",
"TCGA-CV-6962-11A-01D-1913-05",
"TCGA-BA-A6D8-01A-31D-A31M-05",
"TCGA-CV-6934-01A-11D-1913-05",
"TCGA-BA-4076-01A-01D-1433-05",
"TCGA-CV-6951-01A-11D-1913-05",
"TCGA-CV-6935-11A-01D-1913-05",
"TCGA-CV-7434-01A-11D-2130-05",
"TCGA-CV-6961-01A-21D-1913-05",
"TCGA-CR-6474-01A-11D-1871-05",
"TCGA-P3-A5QE-01A-11D-A28S-05",
"TCGA-UF-A7JA-01A-12D-A34K-05",
"TCGA-BB-7871-01A-11D-2230-05",
"TCGA-CV-5976-01A-11D-1684-05",
"TCGA-CR-6488-01A-12D-2079-05",
"TCGA-BA-6868-01B-12D-1913-05",
"TCGA-CV-A6JY-01A-11D-A31M-05",
"TCGA-CV-5439-01A-01D-1684-05",
"TCGA-CV-6936-01A-11D-1913-05",
"TCGA-D6-A6EO-01A-11D-A31M-05",
"TCGA-BA-4077-01B-01D-1433-05",
"TCGA-CQ-6220-01A-11D-1913-05",
"TCGA-CV-5432-11B-01D-1684-05",
"TCGA-CN-6996-01A-11D-1913-05",
"TCGA-CV-6938-01A-11D-1913-05",
"TCGA-P3-A6T5-01A-11D-A34K-05",
"TCGA-BA-A6DI-01A-11D-A30F-05",
"TCGA-CN-A63T-01A-11D-A28S-05",
"TCGA-HD-7754-01A-11D-2079-05",
"TCGA-D6-6515-01A-21D-1871-05",
"TCGA-MZ-A6I9-01A-11D-A31M-05",
"TCGA-CN-A63W-01A-11D-A30F-05",
"TCGA-CV-5430-01A-02D-1684-05",
"TCGA-CN-6011-01A-11D-1684-05",
"TCGA-CQ-A4C7-01A-11D-A254-05",
"TCGA-D6-A6EQ-01A-11D-A31M-05",
"TCGA-CR-6478-01A-11D-1871-05",
"TCGA-CX-7219-01A-11D-2014-05",
"TCGA-P3-A6T8-01A-11D-A34K-05",
"TCGA-CQ-7064-01A-11D-2398-05",
"TCGA-CV-5973-01A-11D-1684-05",
"TCGA-KU-A66S-01A-21D-A30F-05",
"TCGA-BB-7861-01A-11D-2230-05",
"TCGA-DQ-5625-01A-01D-1871-05",
"TCGA-CN-5363-01A-01D-1433-05",
"TCGA-H7-7774-01A-21D-2079-05",
"TCGA-CV-7435-01A-11D-2130-05",
"TCGA-CV-A45Q-01A-11D-A24I-05",
"TCGA-BB-7864-01A-11D-2230-05",
"TCGA-CR-7370-01A-11D-2130-05",
"TCGA-BA-5559-01A-01D-1511-05",
"TCGA-HD-A633-01A-11D-A28S-05",
"TCGA-UF-A7JC-01A-21D-A34K-05",
"TCGA-CV-6955-11A-01D-2014-05",
"TCGA-CN-5360-01A-01D-1433-05",
"TCGA-CN-4733-01A-02D-1871-05",
"TCGA-CR-7385-01A-11D-2014-05",
"TCGA-CV-7242-01A-11D-2014-05",
"TCGA-CV-5440-01A-01D-1511-05",
"TCGA-P3-A6T6-01A-11D-A34K-05",
"TCGA-CV-7247-01A-11D-2014-05",
"TCGA-CR-5249-01A-01D-1511-05",
"TCGA-CV-6941-01A-11D-1913-05",
"TCGA-CV-7410-01A-21D-2079-05",
"TCGA-CV-7250-01A-11D-2014-05",
"TCGA-CN-6012-01A-11D-1684-05",
"TCGA-KU-A6H8-01A-21D-A34K-05",
"TCGA-UF-A71B-01A-12D-A34K-05",
"TCGA-CV-A45Z-01A-21D-A254-05",
"TCGA-CV-6955-01A-11D-2064-05",
"TCGA-D6-A4Z9-01A-11D-A254-05",
"TCGA-CV-6940-01A-11D-1913-05",
"TCGA-WA-A7H4-01A-21D-A34K-05",
"TCGA-UF-A7JK-01A-11D-A34K-05",
"TCGA-CR-6484-01A-11D-1871-05",
"TCGA-CV-5971-11A-01D-1684-05",
"TCGA-CR-7402-01A-11D-2014-05",
"TCGA-F7-A50G-01A-11D-A265-05",
"TCGA-CR-7389-01A-11D-2014-05",
"TCGA-DQ-7590-01A-11D-2230-05",
"TCGA-CV-7180-01A-11D-2014-05",
"TCGA-CV-7101-11A-01D-2014-05",
"TCGA-BB-4227-01A-01D-1871-05",
"TCGA-CV-A45X-01A-21D-A254-05",
"TCGA-UF-A7JV-01A-11D-A34K-05",
"TCGA-CN-5367-01A-01D-1433-05",
"TCGA-CV-7245-11A-01D-2014-05",
"TCGA-KU-A6H7-06A-21D-A31M-05",
"TCGA-QK-A6IJ-01A-11D-A31M-05",
"TCGA-CN-6995-01A-31D-2014-05",
"TCGA-CR-7379-01A-11D-2014-05",
"TCGA-CN-A499-01A-11D-A24I-05",
"TCGA-D6-6824-01A-11D-1913-05",
"TCGA-CV-5442-01A-01D-1511-05",
"TCGA-CR-6470-01A-11D-1871-05",
"TCGA-T2-A6WZ-01A-21D-A34K-05",
"TCGA-CV-6935-01A-11D-1913-05",
"TCGA-CV-A6JN-01A-11D-A31M-05",
"TCGA-UF-A71D-01A-12D-A34K-05",
"TCGA-TN-A7HL-01A-11D-A34K-05",
"TCGA-H7-A6C4-01A-11D-A30F-05",
"TCGA-CV-5970-11A-01D-1684-05",
"TCGA-T2-A6WX-01A-12D-A34K-05",
"TCGA-CV-5434-11B-01D-1684-05",
"TCGA-BA-5149-01A-01D-1511-05",
"TCGA-CR-7392-01A-11D-2014-05",
"TCGA-QK-A8Z7-01A-11D-A392-05",
"TCGA-BB-4225-01A-01D-1433-05",
"TCGA-DQ-5631-01A-01D-1871-05",
"TCGA-UF-A7JO-01A-11D-A34K-05",
"TCGA-D6-A6EM-01A-21D-A31M-05",
"TCGA-QK-A6IG-01A-11D-A31M-05",
"TCGA-H7-8502-01A-11D-2398-05",
"TCGA-CN-4722-01A-01D-1433-05",
"TCGA-CV-7183-01A-11D-2014-05",
"TCGA-CV-A468-01A-11D-A265-05",
"TCGA-MZ-A5BI-01A-31D-A34K-05",
"TCGA-CR-6472-01A-11D-1871-05",
"TCGA-CN-6021-01A-11D-1684-05",
"TCGA-CN-6017-01A-11D-1684-05",
"TCGA-MT-A67G-01A-11D-A30F-05",
"TCGA-BA-6869-01A-11D-1871-05",
"TCGA-D6-A6EP-01A-11D-A31M-05",
"TCGA-BA-A6DJ-01A-11D-A30F-05",
"TCGA-CV-7177-01A-11D-2014-05",
"TCGA-BB-A5HY-01A-11D-A28S-05",
"TCGA-CN-5364-01A-01D-1433-05",
"TCGA-P3-A6T7-01A-11D-A34K-05",
"TCGA-CN-A6V7-01A-12D-A34K-05",
"TCGA-CV-5439-11B-01D-1684-05",
"TCGA-F7-A61W-01A-11D-A28S-05",
"TCGA-CR-7395-01A-11D-2014-05",
"TCGA-CR-6471-01A-11D-1871-05",
"TCGA-QK-A8ZB-01A-11D-A392-05",
"TCGA-D6-A6ES-01A-12D-A31M-05",
"TCGA-CN-A49C-01A-11D-A24I-05",
"TCGA-D6-A6EK-01A-11D-A31M-05",
"TCGA-CN-A63Y-01A-11D-A30F-05",
"TCGA-QK-AA3K-01A-11D-A392-05",
"TCGA-CV-5435-11B-01D-1684-05",
"TCGA-BB-7872-01A-11D-2230-05",
"TCGA-HD-A4C1-01A-11D-A24I-05",
"TCGA-CN-4735-01A-01D-1433-05",
"TCGA-CN-A6UY-01A-12D-A34K-05",
"TCGA-QK-A6IF-01A-11D-A31M-05",
"TCGA-D6-8569-01A-11D-2398-05",
"TCGA-CV-7446-01A-11D-2230-05",
"TCGA-UF-A7J9-01A-12D-A34K-05",
"TCGA-CV-6441-11A-01D-1684-05",
"TCGA-D6-6825-01A-21D-1913-05",
"TCGA-P3-A6SW-01A-11D-A34K-05",
"TCGA-CN-6992-01A-11D-1913-05",
"TCGA-CV-6952-01A-11D-1913-05",
"TCGA-BA-4074-01A-01D-1433-05",
"TCGA-HD-7917-01A-11D-2230-05",
"TCGA-CV-6954-01A-11D-1913-05",
"TCGA-BB-4228-01A-01D-1433-05",
"TCGA-CQ-A4CD-01A-21D-A254-05",
"TCGA-CN-A642-01A-12D-A30F-05",
"TCGA-CQ-6221-01A-11D-2079-05",
"TCGA-CV-6939-01A-11D-1913-05",
"TCGA-KU-A6H7-01A-11D-A31M-05",
"TCGA-CV-7423-01A-11D-2079-05",
"TCGA-CV-7235-01A-11D-2014-05",
"TCGA-CV-7428-01A-11D-2130-05",
"TCGA-D6-6516-01A-11D-1871-05",
"TCGA-CR-7382-01A-11D-2130-05",
"TCGA-BB-4217-01A-11D-2079-05",
"TCGA-HD-7832-01A-11D-2130-05",
"TCGA-CR-7383-01A-11D-2130-05",
"TCGA-CQ-6222-01A-11D-1913-05",
"TCGA-CN-A49B-01A-31D-A24I-05",
"TCGA-CQ-7068-01A-11D-2079-05",
"TCGA-CV-5436-11A-01D-1511-05",
"TCGA-CR-7372-01A-11D-2014-05",
"TCGA-HD-A6HZ-01A-12D-A31M-05",
"TCGA-CV-7407-01A-11D-2079-05",
"TCGA-CV-7416-01A-11D-2079-05",
"TCGA-CN-4727-01A-01D-1433-05",
"TCGA-D6-6823-01A-11D-1913-05",
"TCGA-UP-A6WW-01A-12D-A34K-05",
"TCGA-CR-7388-01A-11D-2014-05",
"TCGA-CN-5358-01A-01D-1511-05",
"TCGA-CQ-7069-01A-11D-2398-05",
"TCGA-CV-6951-11A-01D-1913-05",
"TCGA-IQ-A6SG-01A-12D-A34K-05",
"TCGA-CV-7424-01A-11D-2079-05",
"TCGA-BB-7863-01A-11D-2230-05",
"TCGA-CV-A6JO-01B-11D-A34K-05",
"TCGA-CN-6998-01A-23D-2014-05",
"TCGA-P3-A6T3-01A-11D-A34K-05",
"TCGA-CV-7421-01A-11D-2079-05",
"TCGA-CQ-5333-01A-01D-2398-05",
"TCGA-BA-A6DL-01A-21D-A30F-05",
"TCGA-CV-7422-01A-21D-2079-05",
"TCGA-CQ-5324-01A-01D-1684-05",
"TCGA-IQ-A61I-01A-11D-A30F-05",
"TCGA-CV-6933-11A-01D-1913-05",
"TCGA-CR-7404-01A-11D-2130-05",
"TCGA-CV-A6JM-01A-11D-A31M-05",
"TCGA-CV-6956-11A-01D-2014-05",
"TCGA-CV-5979-11A-01D-1684-05",
"TCGA-HD-8635-01A-11D-2398-05",
"TCGA-CN-6010-01A-11D-1684-05",
"TCGA-CV-6961-11A-01D-1913-05",
"TCGA-CV-A6JZ-01A-11D-A31M-05",
"TCGA-F7-A622-01A-11D-A28S-05",
"TCGA-F7-A50I-01A-11D-A28S-05",
"TCGA-CQ-A4CE-01A-11D-A265-05",
"TCGA-CV-A45W-01A-11D-A254-05",
"TCGA-CN-A63V-01A-11D-A28S-05",
"TCGA-HD-7831-01A-11D-2130-05",
"TCGA-CR-7386-01A-11D-2014-05",
"TCGA-CV-A6K0-01B-21D-A31M-05",
"TCGA-BA-5556-01A-01D-1511-05",
"TCGA-QK-A6V9-01A-11D-A34K-05",
"TCGA-CQ-A4CB-01A-11D-A254-05",
"TCGA-UF-A718-01A-22D-A34K-05",
"TCGA-CX-7085-01A-21D-2014-05",
"TCGA-CV-7418-01A-11D-2079-05",
"TCGA-CV-6959-01A-11D-1913-05",
"TCGA-CV-7252-01A-11D-2014-05",
"TCGA-CV-6948-01A-11D-1913-05",
"TCGA-CN-4726-01A-01D-1433-05",
"TCGA-CV-7101-01A-11D-2014-05",
"TCGA-CV-7089-11A-01D-2014-05",
"TCGA-CQ-6228-01A-11D-1913-05",
"TCGA-CV-7427-01A-11D-2079-05",
"TCGA-CV-7430-01A-11D-2130-05",
"TCGA-CN-5359-01A-01D-1433-05",
"TCGA-P3-A5QA-01A-11D-A28S-05",
"TCGA-HD-8224-01A-11D-2398-05",
"TCGA-BB-8601-01A-11D-2398-05",
"TCGA-CV-7255-11A-01D-2014-05",
"TCGA-CV-7104-01A-11D-2014-05",
"TCGA-CV-6003-01A-11D-1684-05",
"TCGA-F7-8489-01A-31D-2398-05",
"TCGA-TN-A7HI-01A-11D-A34K-05",
"TCGA-CX-7082-01A-11D-2014-05",
"TCGA-CQ-A4C9-01A-11D-A254-05",
"TCGA-DQ-7593-01A-11D-2230-05",
"TCGA-HL-7533-01A-11D-2230-05",
"TCGA-BB-7870-01A-11D-2230-05",
"TCGA-QK-A6II-01A-11D-A31M-05",
"TCGA-CV-6433-01A-11D-1684-05",
"TCGA-CQ-7065-01A-11D-2079-05",
"TCGA-C9-A47Z-01A-11D-A24I-05",
"TCGA-IQ-A61G-01A-11D-A30F-05",
"TCGA-H7-A76A-01A-51D-A34K-05",
"TCGA-CN-6016-01A-11D-1684-05",
"TCGA-UF-A71A-06A-11D-A392-05",
"TCGA-CV-7236-01A-11D-2014-05",
"TCGA-CV-A45U-01A-12D-A24I-05",
"TCGA-BA-5151-01A-01D-1433-05",
"TCGA-CN-4731-01A-01D-1433-05",
"TCGA-CN-5366-01A-01D-1433-05",
"TCGA-CN-5355-01A-01D-1433-05",
"TCGA-CN-4736-01A-01D-1433-05",
"TCGA-UF-A71A-01A-22D-A34K-05",
"TCGA-CN-6994-01A-11D-1913-05",
"TCGA-CV-A6JT-01A-11D-A31M-05",
"TCGA-F7-A624-01A-22D-A30F-05",
"TCGA-CV-5440-11A-01D-1511-05",
"TCGA-UF-A7JT-01A-11D-A34K-05"

)
```


```{r}
query.met.test <- GDCquery(
  project = "TCGA-HNSC", 
  legacy = TRUE,
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450", 
  barcode = barcodeslist
)
GDCdownload(query.met.test)

met.hnsc.450 <- GDCprepare(
  query = query.met.test,
  save = FALSE, 
  # save.filename = "C:/Users/yama16/Desktop/20210927/downloads/hnscDNAmet450k.rda",
  summarizedExperiment = TRUE
)
```


^It runs out of memory when running GDCprepare()

**Can I see which clinical information is available and REDO the query on a subset for which clinical information is available?**


```{r}
query <- GDCquery(project = "TCGA-HNSC", 
                  data.category = "Clinical",
                  data.type = "Clinical Supplement", 
                  data.format = "BCR Biotab")
GDCdownload(query)
clinical.BCRtab.all <- GDCprepare(query)
names(clinical.BCRtab.all)
```

```{r}
View(clinical.BCRtab.all$clinical_nte_hnsc)
View(clinical.BCRtab.all$clinical_patient_hnsc)
View(clinical.BCRtab.all$clinical_follow_up_v1.0_hnsc)
View(clinical.BCRtab.all$clinical_drug_hnsc)
View(clinical.BCRtab.all$clinical_radiation_hnsc)
View(clinical.BCRtab.all$clinical_omf_v4.0_hnsc)
```


**Do this separately**

```{r}
barcodeslist<-c(
  
 "TCGA-UF-A7JV-01A-11D-A34K-05",
"TCGA-CN-5367-01A-01D-1433-05",
"TCGA-CV-7245-11A-01D-2014-05",
"TCGA-KU-A6H7-06A-21D-A31M-05",
"TCGA-QK-A6IJ-01A-11D-A31M-05",
"TCGA-CN-6995-01A-31D-2014-05",
"TCGA-CR-7379-01A-11D-2014-05",
"TCGA-CN-A499-01A-11D-A24I-05",
"TCGA-D6-6824-01A-11D-1913-05",
"TCGA-CV-5442-01A-01D-1511-05",
"TCGA-CR-6470-01A-11D-1871-05",
"TCGA-T2-A6WZ-01A-21D-A34K-05",
"TCGA-CV-6935-01A-11D-1913-05",
"TCGA-CV-A6JN-01A-11D-A31M-05",
"TCGA-UF-A71D-01A-12D-A34K-05",
"TCGA-TN-A7HL-01A-11D-A34K-05",
"TCGA-H7-A6C4-01A-11D-A30F-05",
"TCGA-CV-5970-11A-01D-1684-05",
"TCGA-T2-A6WX-01A-12D-A34K-05",
"TCGA-CV-5434-11B-01D-1684-05",
"TCGA-BA-5149-01A-01D-1511-05",
"TCGA-CR-7392-01A-11D-2014-05",
"TCGA-QK-A8Z7-01A-11D-A392-05",
"TCGA-BB-4225-01A-01D-1433-05",
"TCGA-DQ-5631-01A-01D-1871-05",
"TCGA-UF-A7JO-01A-11D-A34K-05",
"TCGA-D6-A6EM-01A-21D-A31M-05",
"TCGA-QK-A6IG-01A-11D-A31M-05",
"TCGA-H7-8502-01A-11D-2398-05",
"TCGA-CN-4722-01A-01D-1433-05",
"TCGA-CV-7183-01A-11D-2014-05",
"TCGA-CV-A468-01A-11D-A265-05",
"TCGA-MZ-A5BI-01A-31D-A34K-05",
"TCGA-CR-6472-01A-11D-1871-05",
"TCGA-CN-6021-01A-11D-1684-05",
"TCGA-CN-6017-01A-11D-1684-05",
"TCGA-MT-A67G-01A-11D-A30F-05",
"TCGA-BA-6869-01A-11D-1871-05",
"TCGA-D6-A6EP-01A-11D-A31M-05",
"TCGA-BA-A6DJ-01A-11D-A30F-05",
"TCGA-CV-7177-01A-11D-2014-05",
"TCGA-BB-A5HY-01A-11D-A28S-05",
"TCGA-CN-5364-01A-01D-1433-05",
"TCGA-P3-A6T7-01A-11D-A34K-05",
"TCGA-CN-A6V7-01A-12D-A34K-05",
"TCGA-CV-5439-11B-01D-1684-05",
"TCGA-F7-A61W-01A-11D-A28S-05",
"TCGA-CR-7395-01A-11D-2014-05",
"TCGA-CR-6471-01A-11D-1871-05",
"TCGA-QK-A8ZB-01A-11D-A392-05",
"TCGA-D6-A6ES-01A-12D-A31M-05",
"TCGA-CN-A49C-01A-11D-A24I-05",
"TCGA-D6-A6EK-01A-11D-A31M-05",
"TCGA-CN-A63Y-01A-11D-A30F-05",
"TCGA-QK-AA3K-01A-11D-A392-05",
"TCGA-CV-5435-11B-01D-1684-05",
"TCGA-BB-7872-01A-11D-2230-05",
"TCGA-HD-A4C1-01A-11D-A24I-05",
"TCGA-CN-4735-01A-01D-1433-05",
"TCGA-CN-A6UY-01A-12D-A34K-05",
"TCGA-QK-A6IF-01A-11D-A31M-05",
"TCGA-D6-8569-01A-11D-2398-05",
"TCGA-CV-7446-01A-11D-2230-05",
"TCGA-UF-A7J9-01A-12D-A34K-05",
"TCGA-CV-6441-11A-01D-1684-05",
"TCGA-D6-6825-01A-21D-1913-05",
"TCGA-P3-A6SW-01A-11D-A34K-05",
"TCGA-CN-6992-01A-11D-1913-05",
"TCGA-CV-6952-01A-11D-1913-05",
"TCGA-BA-4074-01A-01D-1433-05",
"TCGA-HD-7917-01A-11D-2230-05",
"TCGA-CV-6954-01A-11D-1913-05",
"TCGA-BB-4228-01A-01D-1433-05",
"TCGA-CQ-A4CD-01A-21D-A254-05",
"TCGA-CN-A642-01A-12D-A30F-05",
"TCGA-CQ-6221-01A-11D-2079-05",
"TCGA-CV-6939-01A-11D-1913-05",
"TCGA-KU-A6H7-01A-11D-A31M-05",
"TCGA-CV-7423-01A-11D-2079-05",
"TCGA-CV-7235-01A-11D-2014-05",
"TCGA-CV-7428-01A-11D-2130-05",
"TCGA-D6-6516-01A-11D-1871-05",
"TCGA-CR-7382-01A-11D-2130-05",
"TCGA-BB-4217-01A-11D-2079-05",
"TCGA-HD-7832-01A-11D-2130-05",
"TCGA-CR-7383-01A-11D-2130-05",
"TCGA-CQ-6222-01A-11D-1913-05",
"TCGA-CN-A49B-01A-31D-A24I-05",
"TCGA-CQ-7068-01A-11D-2079-05",
"TCGA-CV-5436-11A-01D-1511-05",
"TCGA-CR-7372-01A-11D-2014-05",
"TCGA-HD-A6HZ-01A-12D-A31M-05",
"TCGA-CV-7407-01A-11D-2079-05",
"TCGA-CV-7416-01A-11D-2079-05",
"TCGA-CN-4727-01A-01D-1433-05",
"TCGA-D6-6823-01A-11D-1913-05",
"TCGA-UP-A6WW-01A-12D-A34K-05",
"TCGA-CR-7388-01A-11D-2014-05",
"TCGA-CN-5358-01A-01D-1511-05",
"TCGA-CQ-7069-01A-11D-2398-05",
"TCGA-CV-6951-11A-01D-1913-05",
"TCGA-IQ-A6SG-01A-12D-A34K-05",
"TCGA-CV-7424-01A-11D-2079-05",
"TCGA-BB-7863-01A-11D-2230-05",
"TCGA-CV-A6JO-01B-11D-A34K-05",
"TCGA-CN-6998-01A-23D-2014-05",
"TCGA-P3-A6T3-01A-11D-A34K-05",
"TCGA-CV-7421-01A-11D-2079-05",
"TCGA-CQ-5333-01A-01D-2398-05",
"TCGA-BA-A6DL-01A-21D-A30F-05",
"TCGA-CV-7422-01A-21D-2079-05",
"TCGA-CQ-5324-01A-01D-1684-05",
"TCGA-IQ-A61I-01A-11D-A30F-05",
"TCGA-CV-6933-11A-01D-1913-05",
"TCGA-CR-7404-01A-11D-2130-05",
"TCGA-CV-A6JM-01A-11D-A31M-05",
"TCGA-CV-6956-11A-01D-2014-05",
"TCGA-CV-5979-11A-01D-1684-05",
"TCGA-HD-8635-01A-11D-2398-05",
"TCGA-CN-6010-01A-11D-1684-05",
"TCGA-CV-6961-11A-01D-1913-05",
"TCGA-CV-A6JZ-01A-11D-A31M-05",
"TCGA-F7-A622-01A-11D-A28S-05",
"TCGA-F7-A50I-01A-11D-A28S-05",
"TCGA-CQ-A4CE-01A-11D-A265-05",
"TCGA-CV-A45W-01A-11D-A254-05",
"TCGA-CN-A63V-01A-11D-A28S-05",
"TCGA-HD-7831-01A-11D-2130-05",
"TCGA-CR-7386-01A-11D-2014-05",
"TCGA-CV-A6K0-01B-21D-A31M-05",
"TCGA-BA-5556-01A-01D-1511-05",
"TCGA-QK-A6V9-01A-11D-A34K-05",
"TCGA-CQ-A4CB-01A-11D-A254-05",
"TCGA-UF-A718-01A-22D-A34K-05",
"TCGA-CX-7085-01A-21D-2014-05",
"TCGA-CV-7418-01A-11D-2079-05",
"TCGA-CV-6959-01A-11D-1913-05",
"TCGA-CV-7252-01A-11D-2014-05",
"TCGA-CV-6948-01A-11D-1913-05",
"TCGA-CN-4726-01A-01D-1433-05",
"TCGA-CV-7101-01A-11D-2014-05",
"TCGA-CV-7089-11A-01D-2014-05",
"TCGA-CQ-6228-01A-11D-1913-05",
"TCGA-CV-7427-01A-11D-2079-05",
"TCGA-CV-7430-01A-11D-2130-05",
"TCGA-CN-5359-01A-01D-1433-05",
"TCGA-P3-A5QA-01A-11D-A28S-05",
"TCGA-HD-8224-01A-11D-2398-05",
"TCGA-BB-8601-01A-11D-2398-05",
"TCGA-CV-7255-11A-01D-2014-05",
"TCGA-CV-7104-01A-11D-2014-05",
"TCGA-CV-6003-01A-11D-1684-05",
"TCGA-F7-8489-01A-31D-2398-05",
"TCGA-TN-A7HI-01A-11D-A34K-05",
"TCGA-CX-7082-01A-11D-2014-05",
"TCGA-CQ-A4C9-01A-11D-A254-05",
"TCGA-DQ-7593-01A-11D-2230-05",
"TCGA-HL-7533-01A-11D-2230-05",
"TCGA-BB-7870-01A-11D-2230-05",
"TCGA-QK-A6II-01A-11D-A31M-05",
"TCGA-CV-6433-01A-11D-1684-05",
"TCGA-CQ-7065-01A-11D-2079-05",
"TCGA-C9-A47Z-01A-11D-A24I-05",
"TCGA-IQ-A61G-01A-11D-A30F-05",
"TCGA-H7-A76A-01A-51D-A34K-05",
"TCGA-CN-6016-01A-11D-1684-05",
"TCGA-UF-A71A-06A-11D-A392-05",
"TCGA-CV-7236-01A-11D-2014-05",
"TCGA-CV-A45U-01A-12D-A24I-05",
"TCGA-BA-5151-01A-01D-1433-05",
"TCGA-CN-4731-01A-01D-1433-05",
"TCGA-CN-5366-01A-01D-1433-05",
"TCGA-CN-5355-01A-01D-1433-05",
"TCGA-CN-4736-01A-01D-1433-05",
"TCGA-UF-A71A-01A-22D-A34K-05",
"TCGA-CN-6994-01A-11D-1913-05",
"TCGA-CV-A6JT-01A-11D-A31M-05",
"TCGA-F7-A624-01A-22D-A30F-05",
"TCGA-CV-5440-11A-01D-1511-05",
"TCGA-UF-A7JT-01A-11D-A34K-05"
  
)

```

```{r}
query.met.test <- GDCquery(
  project = "TCGA-HNSC", 
  legacy = TRUE,
  data.category = "DNA methylation",
  platform = "Illumina Human Methylation 450", 
  barcode = barcodeslist
)
GDCdownload(query.met.test)

met.hnsc.450_5 <- GDCprepare(
  query = query.met.test,
  save = FALSE, 
  save.filename = "C:/Users/yama16/Desktop/20210927/downloads/hnscDNAmet450k_5.rda",
  summarizedExperiment = TRUE
)
```

```{r}
# Save multiple objects
save(hnsc1,hnsc2,hnsc3, hnsc4, hnsc5, file = "C:/Users/yama16/Desktop/20210927/downloads/hnsc_all.RData")
```

```{r}
yert<-cbind(hnsc1,hnsc2,hnsc3, hnsc4, hnsc5)
```

```{r}
save(yert, file = "C:/Users/yama16/Desktop/20210927/downloads/hnsc_merged.rds")
load(file = "C:/Users/yama16/Desktop/20210927/downloads/hnsc_merged.rds")
```



# Cleaning

## Drop CPGs that all have NAs in the row

```{r}
allna<-which(rowSums(yert,na.rm = TRUE)==0)

# yert[allna,] %>% View()
yert_allnadrop<-yert[-allna,]
dim(yert_allnadrop)
```

## Drop CPGs that have greater than X% NAs in the row (as in, 50% or more of patients have "NA" values for these CGs)

```{r}
mycutoff<-0.3
cutoffrowna<-mycutoff*ncol(yert)

numberofna<-which(rowSums(is.na(yert))/cutoffrowna > mycutoff)

yert_cutoffna<-yert[-numberofna,]
dim(yert_cutoffna)
```
^An extra X amount of CPGs dropped after making the cutoff less stringent (if all the row is "NA" vs if greater than X% of the row is NA)


### Impute the rows that have X% or less missing
Mean imputation works best, from another exercise I did before:
```{r}
meanimputation <- function(x) {x[is.na(x)] <- mean(x, na.rm=TRUE); return(x)}
yert_cutoffna_impute  <- as.data.frame(apply(yert_cutoffna, 1, meanimputation ))
yert_cutoffna_impute<-t(yert_cutoffna_impute)
```



### Keep top 1000 most variably expressed CPGs

```{r}
cutoff_top1000<-min(head(sort(rowSds(yert_cutoffna_impute), decreasing=TRUE),1000))
keepers_top1000<-which(rowSds(yert_cutoffna_impute) >= cutoff_top1000)
yert_cutoffna_impute_top1000<-yert_cutoffna_impute[keepers_top1000,]
dim(yert_cutoffna_impute_top1000)
```

```{r}
write.csv(yert_cutoffna_impute_top1000, "hnsc_cpg1000.csv")
```


## Drop all NAs

```{r}
dropthese<-which(is.na(rowSums(yert)))

yertdrop<-yert[-dropthese,]

dim(yertdrop)
```


## Apply standard deviation filter

package matrixStats 
```{r}
# plot mean vs standard deviation
smoothScatter(x=rowMeans2(yertdrop), y=rowSds(yertdrop))

abline(h=0.05, col=2)
abline(v=0.15,col=2)
```

**We could cut these out**

x=rowMeans2(yertdrop), y=rowSds(yertdrop)

```{r}
lowsdmean<-which(rowMeans2(yertdrop)>0.15 & rowSds(yertdrop)>0.05)

yertdrop_sdmean<-yertdrop[lowsdmean,]
dim(yertdrop_sdmean)
```
375044 (all NAs dropped) -> 224501 (with SD and mean cutoff)



