---
title: "stat6500proj_refined"
author: "Kyoko Yamaguchi"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Intro
## Packages
```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(Rtsne)
library(pals)
library(umap)
```

## Read in files
```{r}
sortedcpg<-readRDS(file="hnsc_cpg1000_sorted.RDS")
```


# Identifying patient clusters in methylation array data
## Dimension reduction of cpgs using TSNE
```{r}
set.seed(123)
cpgtsne<-Rtsne(sortedcpg, perplexity=15, exaggeration_factor=5)$Y
```

```{r}
par(pty="s")
plot(cpgtsne, xlab= "TSNE dim 1", ylab="TSNE dim 2")
```

From the above, it appears there are about 5 clusters, maybe 6.

Dimension reduction using UMAP:

```{r}
myumapsetting<-umap.defaults
myumapsetting$spread<-4 #used during automatic estimation of a/b parameters
myumapsetting$min_dist<-0.5 #must be smaller than "spread"; how close points appear in final layout
```

```{r}
# UMAP
set.seed(123)
cpgumap<-umap(sortedcpg, myumapsetting)
par(pty="s")
plot(cpgumap[["layout"]], xlab="UMAP dimension 1", ylab="UMAP dimension 2")
```

## Using the original matrix of CPGs to find patient clusters

```{r}
library(Mercator)
library(clValid)
```


## clValid
clValid can compare internal validation metrics between Kmeans, hierarchical, and pam clustering for different values of K.
```{r}
myclmethods<-c("hierarchical", "kmeans", "pam")
clvalidres<-clValid(sortedcpg, nClust=2:10, clMethods = myclmethods, validation="internal")
```


```{r}
clvalidres@measures
```
```{r}
plot(clvalidres)
```
```{r}
summary(clvalidres)
```

Connectivity should be minimized; Dunn index should be maximized; Silhouette width should be maximized.


# How many K to explore?
```{r}
myK<-seq(from=3, to=6, by=1)
```

## kmeans
```{r}
res_kmeans2<-kmeans(x=as.matrix(sortedcpg), centers=2, iter.max = 10, nstart = 1,
       algorithm = c("Hartigan-Wong"), trace=FALSE)
```

```{r}
mykmeansclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  mykmeansres = kmeans(x=as.matrix(sortedcpg), centers=myK[i], 
                                 iter.max = 10, nstart = 1,
       algorithm = c("Hartigan-Wong"), trace=FALSE)
  mykmeansclusters [,i] <-mykmeansres$cluster
}

colnames(mykmeansclusters)<-paste("Kmeans",myK, sep="_")
```


## kmedoids (PAM)
res_pam2<-pam(x=LFt, k=2, metric = "euclidean", stand = FALSE)
```{r}
mypamclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  mypamres = pam(x=as.matrix(sortedcpg), k=myK[i], 
                 metric = "euclidean", stand = FALSE)
  mypamclusters [,i] <-mypamres$clustering
}

colnames(mypamclusters)<-paste("PAM",myK, sep="_")
```

## kernlab::specc (spectral clustering)
```{r}
library(kernlab)
```

```{r}
myspeccclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  myspeccres<-specc(as.matrix(sortedcpg), centers=myK[i])
  myspeccclusters [,i] <-myspeccres@.Data
}

colnames(myspeccclusters)<-paste("specc",myK, sep="_")
```


## mclust
```{r}
library(mclust)
```

```{r}
# compute BIC for all covariance structures and up to 9 components
# Gives a glimpse as to which model-component pairs are best for our data according to BIC
myG <- myK

mybic<-mclustBIC(data=sortedcpg, G=myG)
mybic
plot(mybic)
```

```{r}
mymodelNames<-c("EII", "VII", "EEI", "VEI", "EVI", "VVI")
matBIC<-as.data.frame(matrix(mybic, ncol=length(mymodelNames), nrow=length(myG)))
colnames(matBIC)<-mymodelNames
rownames(matBIC)<-myG
top<-3
topBIC<-head(sort(mybic, decreasing=TRUE),top)
myrow<-rep(NA, top)
mycol<-rep(NA, top)

for(i in 1:length(topBIC)){
w<-which(topBIC[i]==matBIC, arr.ind = TRUE)
print(w)
myrow[i]<-rownames(matBIC)[w[1,1]]
mycol[i]<-colnames(matBIC)[w[1,2]]
}
```

```{r}
mymclusters<-matrix(NA, nrow=nrow(sortedcpg), ncol=top)
for (i in 1:top){
  resmclust<-Mclust(data=sortedcpg, G=myrow[i], modelNames = mycol[i])
  mymclusters[,i]<-resmclust$classification
}
```

```{r}
colnames(mymclusters)<-paste("mclust", mycol, myrow, sep="_")
head(mymclusters)
```

```{r}
# what if mclust with 2 groups?
resmclust2<-Mclust(data=sortedcpg, G=5, modelNames = "VVI")
```

```{r}
par(pty="s")
  plot(x=cpgtsne[,1], y=cpgtsne[,2], 
     bg=alphabet()[resmclust2$classification], pch=21, cex=1.2,
     xlab="t-SNE dimension 1",ylab="t-SNE dimension 2")
```


# Combining all the clusters tables into 1

```{r}
clustersbig<-cbind(mykmeansclusters, mypamclusters, myspeccclusters, 
                   mymclusters) #just plotting 1
rownames(mykmeansclusters)<-rownames(sortedcpg)
```

# A bunch of plots

```{r, out.width="50%"}
par(pty="s")

for (i in 1:ncol(clustersbig)){
  
  plot(x=cpgtsne[,1], y=cpgtsne[,2], 
     bg=glasbey()[clustersbig[,i]], pch=21, cex=1.2,
     xlab="t-SNE dimension 1",ylab="t-SNE dimension 2",
     main=colnames(clustersbig)[i])
}

```

```{r, out.width="50%"}
par(pty="s")

for (i in 1:ncol(clustersbig)){
  
  plot(x=cpgumap$layout[,1], y=cpgumap$layout[,2], 
     bg=glasbey()[clustersbig[,i]], pch=21, cex=1.2,
     xlab="UMAP dimension 1",ylab="UMAP dimension 2",
     main=colnames(clustersbig)[i])
}
```


## Using Jaccard distance matrix to find patient clusters

# How many K to explore?
```{r}
myK<-seq(from=3, to=6, by=1) #changed this to evaluate only a few
```


## Jaccard distance 
```{r}
library(Mercator)
```


## TSNE/UMAP on Jaccard distance

```{r}
myjacc<-binaryDistance(t(sortedcpg), metric="jaccard")
set.seed(123)
myjacctsne<-Rtsne(myjacc, perplexity=15, exaggeration_factor=10)
mysokal<-binaryDistance(t(sortedcpg), metric="sokalMichener")
set.seed(456)
mysokaltsne<-Rtsne(mysokal, perplexity=15, exaggeration_factor=10)
```


```{r}
par(mfrow=c(1,2), pty="s")
plot(myjacctsne$Y, xlab="t-SNE dimension 1",ylab="t-SNE dimension 2",
     main="TSNE on Jaccard distance")
plot(mysokaltsne$Y, xlab="t-SNE dimension 1",ylab="t-SNE dimension 2",
     main="TSNE on Sokal-Michener distance")
```


# Identifying clusters of patients from CPG matrix

Dimension reduction using UMAP:

```{r}
myumapsetting<-umap.defaults
myumapsetting$spread<-4 #used during automatic estimation of a/b parameters
myumapsetting$min_dist<-0.5 #must be smaller than "spread"; how close points appear in final layout
```

```{r}
# UMAP
set.seed(123)
myjaccumap<-umap(as.matrix(myjacc), myumapsetting)
par(pty="s")
plot(myjaccumap[["layout"]], xlab="UMAP dimension 1", ylab="UMAP dimension 2")
```


## kmeans

```{r}
mykmeansclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  mykmeansres = kmeans(x=myjacc, centers=myK[i], 
                                 iter.max = 10, nstart = 1,
       algorithm = c("Hartigan-Wong"), trace=FALSE)
  mykmeansclusters [,i] <-mykmeansres$cluster
}

colnames(mykmeansclusters)<-paste("Kmeans",myK, sep="_")
```


## kmedoids (PAM)
res_pam2<-pam(x=LFt, k=2, metric = "euclidean", stand = FALSE)
```{r}
mypamclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  mypamres = pam(x=myjacc, k=myK[i], 
                 metric = "euclidean", stand = FALSE)
  mypamclusters [,i] <-mypamres$clustering
}

colnames(mypamclusters)<-paste("PAM",myK, sep="_")
```

## kernlab::specc (spectral clustering)
```{r}
library(kernlab)
```

```{r}
myspeccclusters <-matrix(NA, ncol=length(myK), nrow=nrow(sortedcpg))

for (i in 1:length(myK)){
  myspeccres<-specc(as.matrix(myjacc), centers=myK[i])
  myspeccclusters [,i] <-myspeccres@.Data
}

colnames(myspeccclusters)<-paste("specc",myK, sep="_")
```



## mclust
```{r}
library(mclust)
```

```{r}
# compute BIC for all covariance structures and up to 9 components
# Gives a glimpse as to which model-component pairs are best for our data according to BIC
myG <- myK
mybic<-mclustBIC(data=as.matrix(myjacc), G=myG)
mybic
plot(mybic)
```

```{r}
mymodelNames<-c("EII", "VII", "EEI", "VEI", "EVI", "VVI")
matBIC<-as.data.frame(matrix(mybic, ncol=length(mymodelNames), nrow=length(myG)))
colnames(matBIC)<-mymodelNames
rownames(matBIC)<-myG
top<-3
topBIC<-head(sort(mybic, decreasing=TRUE),top)
myrow<-rep(NA, top)
mycol<-rep(NA, top)

for(i in 1:length(topBIC)){
w<-which(topBIC[i]==matBIC, arr.ind = TRUE)
print(w)
myrow[i]<-rownames(matBIC)[w[1,1]]
mycol[i]<-colnames(matBIC)[w[1,2]]
}
```

```{r}
mymclusters<-matrix(NA, nrow=nrow(sortedcpg), ncol=top)
for (i in 1:top){
  resmclust<-Mclust(data=as.matrix(myjacc), G=myrow[i], modelNames = mycol[i])
  mymclusters[,i]<-resmclust$classification
}
```

```{r}
colnames(mymclusters)<-paste("mclust", mycol, myrow, sep="_")
head(mymclusters)
```



## Combining all the clusters tables into 1

```{r}
clustersbig<-cbind(mykmeansclusters, mypamclusters, myspeccclusters, 
                   mymclusters) #just plotting 1
rownames(mykmeansclusters)<-rownames(sortedcpg)
```

## A bunch of plots


```{r, out.width="50%"}
par(pty="s")

for (i in 1:ncol(clustersbig)){
  plot(x=myjacctsne$Y[,1], y=myjacctsne$Y[,2], 
     bg=glasbey()[clustersbig[,i]], pch=21, cex=1.2,
     xlab="t-SNE dimension 1",ylab="t-SNE dimension 2",
     main=colnames(clustersbig)[i])
}

```

```{r, out.width="50%"}
par(pty="s")

for (i in 1:ncol(clustersbig)){
  
  plot(x=myjaccumap$layout[,1], y=myjaccumap$layout[,2], 
     bg=glasbey()[clustersbig[,i]], pch=21, cex=1.2,
     xlab="UMAP dimension 1",ylab="UMAP dimension 2",
     main=colnames(clustersbig)[i])
}
```

Save cluster calls with from PAM K=3 using Jaccard distance:

```{r}
clustersfinal<-clustersbig[,"PAM_3"]
```

```{r, eval=FALSE}
saveRDS(clustersfinal, "clustersfinal.RDS")
saveRDS(clustersbig,"clustersbig.RDS")
```


# Bringing in the clinical data
## Overlays of the clinical variables (converted to binary) on top of final clusters

```{r}
# read in clinical table
sortedclin<-readRDS(file="hnsc_clin_sorted.RDS")
dim(sortedclin)
```

Identify which columns in sortedclin have ALL NA values and drop them.
```{r}
isna<-apply(sortedclin, 2, function(x) sum(is.na(x))==nrow(sortedclin))
table(isna)
sortedclin<-sortedclin[,-which(isna)] #save over original
```

```{r}
numlevels<-apply(sortedclin, 2, function(x) length(unique(x)))
# sort(numlevels)
names(numlevels)[which(numlevels<=10)]

interesting<-names(numlevels)[which(numlevels<=10)]
subsetclin<-sortedclin[,interesting]
```

```{r, eval=FALSE, include=FALSE}
par(pty="s")
for (i in 1:length(interesting)){
  plot(myjacctsne$Y, col=as.factor(subsetclin[,i]), main=interesting[i])
  legend('topright', legend = levels(as.factor(subsetclin[,i])), col = 1:3, cex = 0.8, pch = 1)
}
```

^We may need to binarize this and plot as shapes on top of the TSNE:

```{r}
library(dummies)
```
```{r}
clinbinary<-dummy.data.frame(subsetclin)
```

 
```{r, out.width="50%"}
# only plot columns that have a substantial amount of information
w<-which(colSums(clinbinary)>=100) #528/3 = 176
clinbinlist <- colnames(clinbinary)[w]
clinbinary2 <-clinbinary[,w]

par(pty="s")
for (i in 1:ncol(clinbinary2)){
  variable<-colnames(clinbinary2)[i]
  plot(myjacctsne$Y, bg=glasbey()[clustersfinal], main=colnames(clinbinary2)[i], pch=21)
  points(myjacctsne$Y[clinbinary2[,i]==1,1],
         myjacctsne$Y[clinbinary2[,i]==1,2],
         lwd=2, col="black", pch=2)
  legend("bottomright", legend=colnames(clinbinary2)[i], 
  col="black", pch=2, cex=0.7)
}

```
  
## Survival analysis using patient clusters and other clinical predictors

```{r, warning=FALSE, message=FALSE}
require(survival)
library(survminer)
library(ggpubr)
library(tibble)
```


### Kaplan-Meier curves for some clinical variables

```{r function}
plot.ggsurv<- function(surv.fit, clin.table) {
  ggsurv<-ggsurvplot(fit =surv.fit, clin.table,
           palette = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7"),
           pval = TRUE, #conf.int = TRUE, 
           legend=c(0.7,0.85),legend.title = "Patient Group",
           xlab = "Time/days", 
           ggtheme = theme_light())
  
  ggsurv$plot <- ggsurv$plot + 
    theme(legend.text = element_text(size = 16, color = "black"),
          axis.title.x =  element_text(size = 16, color = "black"),
          axis.title.y =  element_text(size = 16, color = "black"))
  
  return(ggsurv)
}
```


```{r}
numlevels<-apply(sortedclin, 2, function(x) length(unique(x)))
# sort(numlevels)
names(numlevels)[which(numlevels<=6)]
```

```{r}
interesting<-names(numlevels)[which(numlevels<=6)]
interestingimportant<-c(interesting, "patient.birth_days_to", "patient.death_days_to")
subsetclin<-sortedclin[,interestingimportant]
```

clin_sub<-subsetclin %>% 
  mutate(patient.age_year= round(-as.numeric(patient.birth_days_to)/365, 2),
         patient.event=sortedclin$patient.vital_status %>% as.factor() %>% as.numeric() -1,  #Alive=0,Dead=1
         ) 

```{r}
clin_sub<-subsetclin %>% 
  mutate(patient.age_year= round(-as.numeric(patient.birth_days_to)/365, 2),
         patient.event=sortedclin$patient.vital_status %>% as.factor() %>% as.numeric() -1,  #Alive=0,Dead=1
         patient.survDays=rowSums(sortedclin[, 
                        c("patient.last_contact_days_to","patient.death_days_to")], na.rm=T))

summary(clin_sub$patient.age_year)
```

There is one patient that has an unknown birth date that needs to be dropped:

```{r}
nobirthday<-which(is.na(clin_sub$patient.birth_days_to))
clin_sub1<-clin_sub[-nobirthday,]
```

```{r}
clin_sub1$patient.age_year %>% summary()
```


```{r}
firstq<-summary(clin_sub1$patient.age_year)[2]
secondq<-summary(clin_sub1$patient.age_year)[4]
thirdq<-summary(clin_sub1$patient.age_year)[5]


clin_sub2<-clin_sub1 %>%
    mutate(age_group = ifelse(patient.age_year>=secondq, "Above_median","Below_median"),
           age_group1 = case_when(patient.age_year <=firstq ~ "1st quantile",
                                  patient.age_year >firstq & patient.age_year <=secondq ~ "2nd quantile",
                                  patient.age_year >secondq & patient.age_year<=thirdq ~ "3rd quantile",
                                  patient.age_year > thirdq ~ "4th quantile"))

```


```{r}
fit_age<- survfit(Surv(patient.survDays, patient.event) ~ age_group1, clin_sub2)
plot.ggsurv(fit_age,clin_sub2)
```


## How does the final cluster calls relate to survival?

```{r}
fit_clusters<- survfit(Surv(patient.survDays, patient.event) ~ clustersfinal, clin_sub)
plot.ggsurv(fit_clusters,clin_sub)
```

Things to do:
-multivariate survival model with cluster calls as one of the predictors
-post-hoc analysis to figure out what cluster is the most 
